Frame Competition

### 2022-07-13 pres/Presentation.Rmd
Finish slides for OX|BER

---
title: "Frame Competition"
subtitle: "How News Framing affects Party Competition"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: true
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

@@

## Motivation: Issue Ownership

@@

## Motivation II

- Parties only option to **remain mute** on other parties' issues?
- Framing literature suggests that issue definitions affect voters' policy preferences.
- Some experimental literature that this can affect issue ownership [@Seeberg2019].

@@

### But can framing affect voting behaviour?

@@

@@

1. Issue framing affects issue preferences.
2. Dominant framing of an issue changes public opinion on that issue.
3. Public votes based on proximity to parties' issue positions.

### Issues more or less beneficial dependent on *dominant framing*!

@@

class: center, inverse, middle

# Research Design

@@

@@

## Measuring Frame Dominance

```{r vectors}

@@

>>>>

---
title: "Frame Competition"
subtitle: "How News Framing affects Party Competition"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

@@

## Issue Ownership Theory

```{r IssueOwnership}

data.frame(
  Party = c("A", "A", "B", "B"),
  Issue = c("A's", "B's", "A's", "B's"),
  Votes = c(0.3, 0.1, 0.1, 0.3)
) %>% 
  ggplot(aes(Issue, Votes, col = Party)) +
  geom_point(size = 4) +
  geom_line(aes(group = Party), size = 2) +
  theme_xaringan() +
  xlab("Salient Issue")

## Parties behave accordingly

@@

## Research Question

- Parties only option to **remain mute** on other parties' issues?
- Framing literature suggests that issue definitions affect voters' policy preferences.
- Experimental work shows that framing can affect issue ownership (Seeberg 2019).

@@

### Does framing affect voting behaviour?

@@

class: inverse

@@

<br>

### 1. Issue framing affects issue attitudes.
### 2. Dominant media framing changes public opinion.
### 3. Policy preferences translate into voting behaviour.

## Media Framing $\rightarrow$ Voting Behaviour

@@

## Research Design

<br>

### 1. Measure similarity of media framing to party framing.
### 2. Estimate effect of similarity on electoral performance.

@@

## Data

### Media Content

Collected **2.5M news articles** from major German newspapers' websites (2013-2019).

.pull-left[

### (Expected) Electoral Performance

]

.pull-right[

  ![](https://www.politico.eu/wp-content/uploads/2022/02/01/Poll-of-Polls.svg)

]

- Polling data collected by *Politico's [Poll of Polls](https://www.politico.eu/europe-poll-of-polls/)*.
- Generates daily moving average of party support.

@@

## Data II

<br>

### Party framing

- 2017 Manifesto corpora, reduced to migration content.
- Ideal because **summarise general position** on the issue.
- Downside: assumption of stable position over time.

## Measuring Issue Framing

.pull-left[

  ![](vis/CognitiveStorage.png)

]

.pull-right[

1. Identify media content relating to issue:
  - Annotate sample
  - Supervised ML (BERT)
2. Identify considerations:
  -  Topic model

]

## Measuring Frame Similarity

```{r vectors, fig.height=7, cache=FALSE}

@@

## Measuring Frame Similarity

### Why frame *similarity*?

- Parties communicate a certain framing either as cause or outcome of their positions.
- If media framing is **similar**, this supports a parties' position.
- As a result, citizens are more likely to vote for party.

### Measurement

$$Similarity_{Party, t} = cos(TopicVector_{Media}, TopicVector_{Party})_t$$

## SVAR Time Series Model

<br>

- VAR allows modeling multiple time series as functions of own and other variables past.
- Structural VAR allows modeling interaction $Performance = Salience*Similarity$.
- Estimate of interest: effect of similarity on electoral performance.

## Instrumental Variable Approach

.left-column[

```{r dagInstrument, fig.align='center'}
DiagrammeR::grViz("
digraph boxes_and_circles {

  rankdir = 'TB'
  bgcolor = '#bfbfbf'

  # a 'graph' statement
  graph [overlap = true, 
          fontsize = 16]

  # several 'node' statements
  node [shape = plaintext,
        fontname = Verdana,
        fontcolor = black]
  Z [label = 'Events', fontcolor = '#004A95']
  X [label = 'Media Framing']
  Y [label = 'Voting']

  # several 'edge' statements
  Z->X [color = '#004A95']
  X->Y [color = black]
}
")

]

.right-column[

Events drive media attention to considerations, e.g.

- Migrant crime,
- Terrorist Attacks,
- Drownings.

### These should only affect public opinion through media coverage.

]

## Instrumental Variable Approach II

<br>

- Framing can be instrumented using exogenous events.
- Predicted frame attention can then be used to estimate causal effect (**2SLS**).

## Summary

<br>

- **Goal**: Project tries to challenge assumption that issues always affect voters similarly.
- **Argument**: dominant media framing of issues affects party support.
- **Methods**: 
  + innovative measure of frame similarity,
  + combination of time-series and causal inference tools (SVAR + IV).

## Open Questions

- Does theory make sense?
- Time Frame (2017 campaign or longer?)
- Manifestos for party frames?
- SVAR correct model? (might go for unidirectional model with 2SLS)
- Additional Experiment?
- ...?

class: inverse, middle, center

# Fin

## Thank you!

.pull-left[![](vis/banner_new.png)]

.pull-right[![](vis/dfg_logo_englisch_blau_en.jpg)]
